% main.tex — Fee Concentration Insurance Specification
% Master document: \input{} for all components

\documentclass[11pt,a4paper]{article}

\input{preamble}

\title{Fee Concentration Insurance\\[4pt]
       \large Mathematical Specification}
\author{ThetaSwap}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We specify an insurance mechanism against adverse fee concentration in
Uniswap V4 liquidity provision. The fee concentration index
$\AT = (\sum_k \thetapos_k \cdot x_k^2)^{1/2}$ measures realized fee
inequality across LP positions. The Ma--Crapis competitive null
$\ATnull = \sqrt{\ThetaSum/\PosCount^2}$ provides the equal-share baseline.
The concentration deviation $\DeltaPlus = \max(0,\, \AT - \ATnull)$
captures excess concentration above the competitive equilibrium and maps to
price space via the odds ratio $p = \DeltaPlus/(1-\DeltaPlus)$.

Econometric identification (quadratic deviation exit hazard model) establishes
an inverted-U relationship: below the turning point $\DeltaStar \approx 0.09$,
concentration is protective (shelter effect); above it, concentration drives
LP exits (Capponi regime). This turning point calibrates insurance demand.

The hook tracks three co-primary state variables---$\AT$, $\ThetaSum$, $\PosCount$---with
$\DeltaPlus$ derived on the fly. The mechanism is perpetual with jump-adjusted
funding and USDC numeraire. The specific protection architecture (CFMM pool,
hook-based mutual insurance, or hybrid) is pending empirical analysis of
fee concentration severity across pool categories.
\end{abstract}

\tableofcontents

\input{payoff}
\input{funding-rate}
\input{invariants}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
