% main.tex — Fee Concentration Insurance Specification
% Master document: \input{} for all components

\documentclass[11pt,a4paper]{article}

\input{preamble}

\title{Fee Concentration Insurance\\[4pt]
       \large Mathematical Specification}
\author{ThetaSwap}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We specify an insurance mechanism against adverse fee concentration in
Uniswap V4 liquidity provision. The fee concentration index
$\AT = (\sum_k \thetapos_k \cdot x_k^2)^{1/2}$ measures realized fee
inequality across LP positions. The Ma--Crapis competitive null
$\ATnull = \sqrt{\ThetaSum/\PosCount^2}$ provides the equal-share baseline.
The concentration deviation $\DeltaPlus = \max(0,\, \AT - \ATnull)$
captures excess concentration above the competitive equilibrium and maps to
price space via the odds ratio $p = \DeltaPlus/(1-\DeltaPlus)$.

Econometric identification (quadratic deviation exit hazard model) establishes
an inverted-U relationship: for the pool we analyzed (ETH/USDC 30bps),
below the turning point $\DeltaStar \approx 0.09$, concentration is protective
(shelter effect); above it, concentration drives LP
exits\footnote{Capponi, Jia \& Zhu (2024) show that JIT liquidity
mechanically concentrates fees in short-lived positions, diluting passive LPs'
effective fee rate and pushing them toward their optimal exit threshold.}.
We find contingent empirical demand from passive liquidity providers for
insurance against adverse competition, and that the optimal payoff function
is quadratic. The turning point $\DeltaStar$ calibrates insurance demand.

The hook tracks three co-primary state variables---$\AT$ (fee concentration
index), $\ThetaSum$ (aggregate turnover rate $\sum_k \thetapos_k$), and
$\PosCount$ (active position count)---with $\DeltaPlus$ derived on the fly.
The mechanism is perpetual with jump-adjusted funding and USDC numeraire.
\end{abstract}

\tableofcontents

\input{payoff}
\input{econometrics}
\input{trading-function}
\input{reserves}
\input{borrow-rate}
\input{funding-rate}
\input{initialization}
\input{invariants}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}
